---
title: "Assignment1 - Exploratory Data Analysis with R"
author: "Geedhu Kizhakepura Velayudhan - Student_ID 8899510"
date: "2023-06-04"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=5, fig.height=5,
                      fig.path='Figs/', echo = TRUE)
```

## PART A

### QUESTION 1

(4 points) You are working Streaming Service. The following statement is made by 
your manager. Based on the examples and discussion in Lecture 1, transform it into a 
question that can be answered with data analytics. Make sure you discuss the logic and 
reasoning you use to transform it and what problems you might define. (NOTE – This 
question is worth 4 marks, so answer appropriately). We have more customers than 
before, but our new customers are streaming less than before

### Answer:

Logic and Reasoning:  

To transform the manager's statement into questions, we need to focus on two elements mentioned:

1. the increase in the number of customers.  

2. the decrease in streaming activity among new customers.

Problems to define:

1. What do you think is the possible cause of less streaming among new customers?

2. From when did you started noticing about the "streaming less activity" among new customers.

3. Have we established clear criteria for categorizing customers as either new or existing? 
For example, we could define "new customers" as those who signed up within the past three months and "existing customers" as those who have been subscribed for more than three months.

4. Did you collect any data about customer's streaming activities of both new and existing customers like time stamps of when customers start and stop streaming, the titles they watch, and the duration of each streaming session.If any already collected may we have the access to the data?

5. By what percentage would you expect the streaming activity to be increased?

### QUESTION 2

2. (4 points) Consider the following three arrays of data. Each array is data for one file sharing 
site. The numbers in the array represent the number of downloads for each site in a day (for 
example, Site A had 28 downloads on the first day, 29 on the second and so on). 
Site A: (28 29 31 28 30 30 30 32 28 33) 
Site B: (23 19 23 33 32 27 20 24 42 32) 
Site C: (27 26 28 25 27 27 30 30 28 26) 
Based on the data provided, and using the skills learned in this class, answer the following 
questions. Make sure to provide evidence for your answers. 
a) (2 points) Which site has the least downloads on a typical day? 
b) (2 points) Which site has the most inconsistent usage? 

### Answer:

```{r}
#Assigning variable to each site

sitea<- c(28, 29, 31, 28, 30, 30, 30, 32, 28, 33)
siteb<- c(23, 19, 23, 33, 32, 27, 20, 24, 42, 32)
sitec<- c(27, 26, 28, 25, 27, 27, 30, 30, 28, 26)


# a)
# By calculating mean of each site and then comparing them will provide the site has 
# the site that has the least downloads on a typical day.

meanA<- mean(sitea)
meanB<- mean(siteb)
meanC<- mean(sitec)

medianA <- median(sitea)
medianB <- median(siteb)
medianC <- median(sitec)

cat("Means:",meanA,meanB,meanC)
cat("Medians:",medianA,medianB,medianC)

# As Mean of Site C is the least, the Site C has the least downloads on a typical day.

# b)
# By calculating the variance of each site, we will get to know which site has the most inconsistent usage.

varA<- var(sitea)
varB<- var(siteb)
varC<- var(sitec)

#As Variance of Site B is the highest, the Site B has the most inconsistent usage.
```

a) (2 points) Which site has the least downloads on a typical day? 

Answer: 
Considering both the mean and the median, we can see that Site B consistently has lower mean and median values compared to Sites A and C. This indicates that on a typical day, Site B has the fewest downloads. While Site C has the lowest mean value, its median is slightly higher than that of Site B, which suggests that Site B generally has fewer downloads in the middle of the range.

b)  (2 points) Which site has the most inconsistent usage?

Answer: As Variance of Site B is the highest, the Site B has the most inconsistent usage.

## PART 2

### 1. BASIC MANIPULATION


```{r}
###BASIC SETUP
# Clean workspace
rm(list=ls())
options(scipen=9)

getwd()
# 1. Read data and converted to dataframe names 'data'

data <- read.table(file = "Assign01.txt", header = TRUE, sep = ",")

# 2. Append your initials to all variables in the data frame

new_names <- paste0(names(data), "_kv")
colnames(data) <- new_names

#Check if initials are added or not
colnames(data)

#Unserstand the structure if data
str(data)

#3. Change each character variable to a factor variable 
data$Manufacturer_kv<- factor(data$Manufacturer_kv)
data$Server_kv <- factor(data$Server_kv)
data$DC_kv<-factor(data$DC_kv)

# to check char typr of data converted to factor or not?
str(data)

#4. Dimensions of the dataset (rows and columns)? 

dimensions <- dim(data)
num_rows<- dimensions[1]
num_cols<- dimensions[2]
#Print dimensions,rows and columns
cat("Dimensions of dataset: ",dimensions)
cat("Number of Rows: ",num_rows)
cat("Number of Columns: ",num_cols)
```

### 2. SUMMARIZING DATA

#### 1.Means and Standard Deviations 
```{r}
# a. Calculate the mean and standard deviation for Server Message Blocks Received.

#data$SMBR_kv<- as.numeric(data$SMBR_kv)
Mean_SMBR<- mean(data$SMBR_kv)
var_SMBR<- var(data$SMBR_kv)
sd_SMBR<- sd(data$SMBR_kv)

#Print mean, variance and standard deviation of SMBR

cat("Mean of SMBR: ",Mean_SMBR)
cat("Variance of SMBR: ",var_SMBR)
cat("Standard deviation of SMBR: ",sd_SMBR)

# b. Use the results above to calculate the coefficient of variation (rounded to 3 decimal places). 

cv_SMBR<- round((sd_SMBR/Mean_SMBR),3)

cat("Coefficient of Variation of SMBR: ",cv_SMBR)

# c. Calculate the mean and standard deviation for Server Message Blocks Transmitted. 

#data$SMBT_kv<- as.numeric(data$SMBT_kv)
mean_SMBT<- mean(data$SMBT_kv)
var_SMBT<- var(data$SMBT_kv)
sd_SMBT<- sd(data$SMBT_kv)

cat("Mean of SMBT: ",mean_SMBT)
cat("Variance of SMBT: ",var_SMBT)
cat("Standard deviation of SMBT: ",sd_SMBT)
# d. Also calculate the coefficient of variation (rounded to 3 decimal places). 
cv_SMBT<- round((sd_SMBT/mean_SMBT),3)

cat("Coefficient of Variation of SMBT is ",cv_SMBT)
```
e. Does the SMBT or SMBR have more variation? 

Answer : SMBT has more variation as Coefficient of Variation is higher for SMBT than SMBR.


#### 2. Calculate the 45th percentile of the number of Server Message Blocks TransmiSed. This calculation should be rounded to the nearest whole number (no decimal places). 

```{r}
Percentile_SMBT <- quantile(data$SMBT_kv,probs=0.45)

cat("45th percentile is: ",Percentile_SMBT)
```

### ORGANIZING DATA

#### 1.Summary Table 

a) Create a table showing the average Server Message Blocks Transmitted by Manufacturer. This should be rounded to two decimal places. 

```{r}

# Creating the summary table
summary_table <- aggregate( data$SMBT_kv ~ data$Manufacturer_kv ,FUN = function(x) round(mean(x), 2))

# Naming the columns
colnames(summary_table) <- c("Manufacturer", "Average SMBT")

#Print the summary table
summary_table

```

b) Which Manufacturer’s Servers have, on average, transmitted the most server message blocks?

Answer: Lled is the Manufacturer who transmitted the most server message blocks.

#### 2.Cross Tabulation

a) Create a table counting all Servers by Data Center.

```{r}
#Create the cross table
cross_table <- table(data$DC_kv,data$Server_kv)
# Print the cross table
cross_table
```

b) Change the table to show the percentage of each Server in each Data Center.This should be rounded to three decimal places.

```{r}
#Create Percentable table
percentage_table <- round(prop.table(cross_table, margin = 2) * 100,3)
# Print Percentage table
percentage_table
```

c) What percentage of servers at Elmira are MG9696?
```{r}
servers_at_elmira <- percentage_table["MG9696","Elmira"]

cat("The percentage of servers at Elmira are MG9696 is",servers_at_elmira)
```
#### 3.Histogram

a) Create a histogram of Server Message Blocks Transmitted. 
b) The plot should be properly labelled and a unique colour and have 10 breaks. 
 For both a and b , below is the code:
 
```{r}
hist(data$SMBT_kv,
     xlab = "Server Message Blocks Transmitted",
     ylab = "Frequency", 
     main = "Histogram of Server Message Blocks Transmitted",
     col = "purple",
     breaks = 10)
```


c) Which range of SMBT is the most common?

Answer:  The most common range in SMBT is 50000 to 60000.
